Table 1. Candidate genes and differentially expressed genes

### 2020-08-19 analysis/boneyard.Rmd
now with summary of DEGs characteriztion

>>>>

## Candidate Correlations - SUppl fig 1

```{r correlations}
hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
gon1 <- makecorrdf("female", "gonad", litNotGOknown)

hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
gon2 <- makecorrdf("male", "gonad", litNotGOknown)

hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
pit3 <- makecorrdf("female", "pituitary", GOgenes)  
gon3 <- makecorrdf("female", "gonad", GOgenes)

hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
pit4 <- makecorrdf("male", "pituitary", GOgenes)  
gon4 <- makecorrdf("male", "gonad", GOgenes)  
```

```{r supplfig-2, fig.width=7.25, fig.height=7.25}

b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
  theme(axis.text.x = element_blank())
b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))

b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
  scale_x_discrete(position = "top") 
b5 <- plotcorrplot(pit2, NULL)    + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
b8 <- plotcorrplot(pit3, NULL)   + 
  theme(axis.text.x = element_blank())
b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))

b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
b11 <- plotcorrplot(pit4, NULL)    + 
  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
mylegend <- get_legend(forlegend)

allcorplots <- plot_grid(b1,b4,b7,b10,
          b2,b5,b8,b11,
          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))

supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
supplfig2

## Figure

```{r fig2, fig.width = 7.25, fig.height = 7.25}

c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
c5 <- scattercorrelations(FG, FG$PGR, "PGR",  FG$ESR1,  "ESR1", "#969696")   + labs(title = " ", subtitle = " ")

c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = " " )  
c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " ) 
```

```{r newfig}

plot3candidates <- function(hypgene, pitgene, gongene){

  a1 <- newcandidateboxplot("hypothalamus", hypgene, "female", "female hypothalamus", "sequential") 
  a2 <- externalboxplots("hypothalamus", hypgene, "female", " ", "external") 
  a3 <- newcandidateboxplot("hypothalamus", hypgene, "male", "male hypothalamus", "sequential")
  a4 <- externalboxplots( "hypothalamus", hypgene, "female", " ", "external")

  hypplots <- plot_grid(a1,a2, a3, a4, rel_widths = c(2,1,2,1), nrow = 1)

  b1 <- newcandidateboxplot("pituitary", pitgene, "female", "female pituitary", "sequential")
  b2 <- externalboxplots("pituitary", pitgene, "female", " ", "external") 
  b3 <- newcandidateboxplot("pituitary", pitgene, "male", "male pituitary", "sequential")
  b4 <- externalboxplots( "pituitary", pitgene, "female", " ", "external")

  pitplots <- plot_grid(b1,b2, b3, b4, rel_widths = c(2,1,2,1), nrow = 1)

  c1 <- newcandidateboxplot("gonad", gongene, "female", "female gonad", "sequential") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c2 <- externalboxplots("gonad", gongene, "female", " ", "external") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c3 <- newcandidateboxplot("gonad", gongene, "male", "male gonad", "sequential") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c4 <- externalboxplots( "gonad", gongene, "female", " ", "external") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  gonplots <- plot_grid(c1,c2, c3, c4, rel_widths = c(2,1,2,1), nrow = 1)

  fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
                       labels = "AUTO", label_size = 8,
                       rel_heights = c(1,1,1.2))
  print(fig2alt)

}

plot3candidates("COMT","PRL","AVPR1A")
plot3candidates("PRL","PRL", "PRL")
plot3candidates("PRL","FSH", "ESR1")
```

### 2020-08-19 analysis/figure-2.Rmd
now with summary of DEGs characteriztion

---
title: "Fig2: Candidate genes"
output: md_document
---

# Candidate gene analysis

@@

## Candidate DEGs

```{r DEGs}

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, lay_inc.d3:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
(table1)
```

## Candidate VSDs

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels) %>%
  mutate(treatment = factor(treatment, levels = charlevels)) %>%
  drop_na() %>%
    mutate(external = fct_collapse(treatment, 
                                   "none" = c("control", "bldg"),
                                   "eggs" = c("lay" , "inc.d3", 
                                              "inc.d9", "inc.d17"),
                                   "chicks" = c("hatch", "n5", "n9")))
head(candidatevsd)

candidatevsdwide <- candidatevsd  %>%
  select(-external) %>%
  pivot_wider(names_from = gene, values_from = counts) 
FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad") 
```

## Candidate Correlations - SUppl fig 1

```{r correlations}
hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
gon1 <- makecorrdf("female", "gonad", litNotGOknown)

hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
gon2 <- makecorrdf("male", "gonad", litNotGOknown)

hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
pit3 <- makecorrdf("female", "pituitary", GOgenes)  
gon3 <- makecorrdf("female", "gonad", GOgenes)

hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
pit4 <- makecorrdf("male", "pituitary", GOgenes)  
gon4 <- makecorrdf("male", "gonad", GOgenes)  
```

```{r supplfig-2, fig.width=7.25, fig.height=7.25}

b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
  theme(axis.text.x = element_blank())
b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))

b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
  scale_x_discrete(position = "top") 
b5 <- plotcorrplot(pit2, NULL)    + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
b8 <- plotcorrplot(pit3, NULL)   + 
  theme(axis.text.x = element_blank())
b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))

b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
b11 <- plotcorrplot(pit4, NULL)    + 
  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
mylegend <- get_legend(forlegend)

allcorplots <- plot_grid(b1,b4,b7,b10,
          b2,b5,b8,b11,
          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))

supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
supplfig2

@@

## Figure

```{r fig2, fig.width = 7.25, fig.height = 7.25}

c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
c5 <- scattercorrelations(FG, FG$PGR, "PGR",  FG$ESR1,  "ESR1", "#969696")   + labs(title = " ", subtitle = " ")

c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = " " )  
c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " ) 
```

```{r newfig}

plot3candidates <- function(hypgene, pitgene, gongene){

  a1 <- newcandidateboxplot("hypothalamus", hypgene, "female", "female hypothalamus", "sequential") 
  a2 <- externalboxplots("hypothalamus", hypgene, "female", " ", "external") 
  a3 <- newcandidateboxplot("hypothalamus", hypgene, "male", "male hypothalamus", "sequential")
  a4 <- externalboxplots( "hypothalamus", hypgene, "female", " ", "external")

  hypplots <- plot_grid(a1,a2, a3, a4, rel_widths = c(2,1,2,1), nrow = 1)

  b1 <- newcandidateboxplot("pituitary", pitgene, "female", "female pituitary", "sequential")
  b2 <- externalboxplots("pituitary", pitgene, "female", " ", "external") 
  b3 <- newcandidateboxplot("pituitary", pitgene, "male", "male pituitary", "sequential")
  b4 <- externalboxplots( "pituitary", pitgene, "female", " ", "external")

  pitplots <- plot_grid(b1,b2, b3, b4, rel_widths = c(2,1,2,1), nrow = 1)

  c1 <- newcandidateboxplot("gonad", gongene, "female", "female gonad", "sequential") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c2 <- externalboxplots("gonad", gongene, "female", " ", "external") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c3 <- newcandidateboxplot("gonad", gongene, "male", "male gonad", "sequential") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  c4 <- externalboxplots( "gonad", gongene, "female", " ", "external") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  gonplots <- plot_grid(c1,c2, c3, c4, rel_widths = c(2,1,2,1), nrow = 1)

  fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
                       labels = "AUTO", label_size = 8,
                       rel_heights = c(1,1,1.2))
  print(fig2alt)

plot3candidates("COMT","PRL","AVPR1A")
plot3candidates("PRL","PRL", "PRL")
plot3candidates("PRL","FSH", "ESR1")
```

```{r writefiles}
#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
#write.csv(candidatevsd, "../results/candidatecounts.csv")
write.csv(table1, "../results/table1-v2.csv")

pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
plot(fig2)
dev.off()

pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
plot(supplfig2)
dev.off()

pdf(file="../figures/fig2-2.pdf", width=7.25, height=7.25)
plot(newfig2)
dev.off()
```

>>>>

---
title: "Table 1. Candidate genes and differentially expressed genes"
output: md_document
---

# Table 1. Candidate genes and differentially expressed genes

@@

## Differentially expressed genes (DEGs)

```{r}
# summary DEG results from DESeq2
allDEGS <- read_csv("../results/suppltable1.csv") %>%
   mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G"),
         comparison = recode(comparison, 
                         "bldg_lay" = "Bldg2Lay",
                         "lay_inc.d3" = "Lay2Inc3", 
                         "inc.d3_inc.d9" = "Inc3toInc9",
                         "inc.d9_inc.d17" = "Inc9toInc17",
                         "inc.d17_hatch" = "In17toHatch",
                         "hatch_n5" = "Hatch2N5",
                         "n5_n9" = "N5toN9"),
         group = paste(sex, tissue, sep = "")) %>%
  mutate(res = paste("(", posneg, ")", sep = "")) %>%
  mutate(compres = paste(comparison, res, sep = "")) %>%
  select(gene, group, compres) 
allDEGS

DEGsbytissue <- function(whichtissueshorthand, whichsex){

  df <- allDEGS %>%
  filter(grepl(whichtissueshorthand, group),
         grepl(whichsex,group)) %>%
  group_by(group,gene) %>%
  summarize(results = str_c(compres, collapse = "; ")) %>%
  mutate(n = (str_count(results, pattern = ";"))+1) %>%
  arrange(desc(n))  %>%
  filter(n>1)
  #group_by(comparisons) %>%
  #summarize(genes = str_c(gene, collapse = "; "))

  print(head(df))
  return(df)

hmDEGs <- DEGsbytissue("H", "M")
pmDEGs <- DEGsbytissue("P", "M")
gmDEGs <- DEGsbytissue("G", "M")

hfDEGs <- DEGsbytissue("H", "F")
pfDEGs <- DEGsbytissue("P", "F")
gfDEGs <- DEGsbytissue("G", "F")

maleDEGs <- rbind(hmDEGs, pmDEGs) %>%
  rbind(., gmDEGs)
maleDEGs

femaleDEGs <- rbind(hfDEGs, pfDEGs) %>%
  rbind(., gfDEGs)
femaleDEGs

summarizedDEGs <- rbind(femaleDEGs,maleDEGs) %>%
  group_by(group, results) %>%
  summarize(genes = str_c(gene, collapse = "; ")) 
summarizedDEGs

@@

```{r writefiles}
write.csv(summarizedDEGs,"../results/summarizedDEGsv2.csv")
